---
title: "New Charter Districts"
output: html_document
---

The following matrix shows the number of newly opened charter school districts by each year between 2004 and 2014 and the number of them that survived in the folowing years. For example, in school year 2014, 310 new charter districts were established, of which 100 were closed and 210 remained operating in 2014. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
data <- read.csv("C:/Users/Amlan/Documents/Amlan's Website/NEW CHARTER DISTRICTS.csv", header = T)
kable(data,  align = c("c", "r"))
```

The following plots show the geographic trends in the openings of new school districts over the past ten years. The plots only show the districts that survived until 2014. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(plotly)
library(dplyr)

df <- read.csv("C:/Users/Amlan/Documents/Amlan's Website/lea_charter_1.csv")
df <- df[order(df$SHYEAR), ]

df$YEAR <- NA
df$YEAR[df$SHYEAR =='2002-03'] <- 2003
df$YEAR[df$SHYEAR =='2003-04'] <- 2004
df$YEAR[df$SHYEAR =='2004-05'] <- 2005
df$YEAR[df$SHYEAR =='2005-06'] <- 2006
df$YEAR[df$SHYEAR =='2006-07'] <- 2007
df$YEAR[df$SHYEAR =='2007-08'] <- 2008
df$YEAR[df$SHYEAR =='2008-09'] <- 2009
df$YEAR[df$SHYEAR =='2009-10'] <- 2010
df$YEAR[df$SHYEAR =='2010-11'] <- 2011
df$YEAR[df$SHYEAR =='2011-12'] <- 2012
df$YEAR[df$SHYEAR =='2012-13'] <- 2013
df$YEAR[df$SHYEAR =='2013-14'] <- 2014

df <- subset(df, df$YEAR>2003)

# common map properties
g <- list(
  scope = 'usa',
  showland = T,
  landcolor = toRGB("gray90"),
  showcountries = F,
  subunitcolor = toRGB("white")
)

one_map <- function(dat) {
  plot_geo(dat) %>%
    add_markers(x = ~LON, y = ~LAT, color = I("blue"), alpha = 0.5) %>%
    add_text(x = -78, y = 47, text = ~unique(YEAR), color = I("black")) %>%
    layout(geo = g)
}

p <- df %>%
  group_by(YEAR) %>%
  do(map = one_map(.)) %>%
  subplot(nrows = 4) %>%
  layout(
    showlegend = FALSE,
    title = 'New Charter Districts per year 2004-2014<br> Source: <a href="https://nces.ed.gov/ccd/pubagency.asp">NCES</a>',
    width = 1000,
    height = 900,
    hovermode = FALSE
  )

htmltools::tagList(list(as.widget(p)))

```
<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- read.csv("C:/Users/Amlan/Documents/Amlan's Website/lea_charter_2.csv")
df <- df[order(df$SHYEAR), ]

df$YEAR <- NA
df$YEAR[df$SHYEAR =='2002-03'] <- 2003
df$YEAR[df$SHYEAR =='2003-04'] <- 2004
df$YEAR[df$SHYEAR =='2004-05'] <- 2005
df$YEAR[df$SHYEAR =='2005-06'] <- 2006
df$YEAR[df$SHYEAR =='2006-07'] <- 2007
df$YEAR[df$SHYEAR =='2007-08'] <- 2008
# df$YEAR[df$SHYEAR =='2008-09'] <- 2009
df$YEAR[df$SHYEAR =='2009-10'] <- 2010
df$YEAR[df$SHYEAR =='2010-11'] <- 2011
df$YEAR[df$SHYEAR =='2011-12'] <- 2012
df$YEAR[df$SHYEAR =='2012-13'] <- 2013
df$YEAR[df$SHYEAR =='2013-14'] <- 2014

df <- subset(df, df$YEAR>2003)

barplot2 <- ggplot(data=df,aes(x=factor(YEAR))) +
    geom_bar(fill="tomato") +
    geom_text(stat='count',aes(label=..count..),vjust=-2, size=3.5) +
    theme_minimal()+
    theme(axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    ggtitle("PLOT 2 - New Mixed Charter Districts per year 2004-2014")

ggplotly(barplot2)
```
<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# common map properties
g <- list(
  scope = 'usa',
  showland = T,
  landcolor = toRGB("gray90"),
  showcountries = F,
  subunitcolor = toRGB("white")
)

one_map <- function(dat) {
  plot_geo(dat) %>%
    add_markers(x = ~LON, y = ~LAT, color = I("red"), alpha = 0.5) %>%
    add_text(x = -78, y = 47, text = ~unique(YEAR), color = I("black")) %>%
    layout(geo = g)
}

df %>%
  group_by(YEAR) %>%
  do(map = one_map(.)) %>%
  subplot(nrows = 4) %>%
  layout(
    showlegend = FALSE,
    title = 'New Mixed Charter Districts per year 2004-2014<br> Source: <a href="https://nces.ed.gov/ccd/pubagency.asp">NCES</a>',
    width = 1000,
    height = 900,
    hovermode = FALSE
  )

```
<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- read.csv("C:/Users/Amlan/Documents/Amlan's Website/lea_charter_2.csv")
df <- df[order(df$SHYEAR), ]

df$YEAR <- NA
df$YEAR[df$SHYEAR =='2002-03'] <- 2003
df$YEAR[df$SHYEAR =='2003-04'] <- 2004
df$YEAR[df$SHYEAR =='2004-05'] <- 2005
df$YEAR[df$SHYEAR =='2005-06'] <- 2006
df$YEAR[df$SHYEAR =='2006-07'] <- 2007
df$YEAR[df$SHYEAR =='2007-08'] <- 2008
# df$YEAR[df$SHYEAR =='2008-09'] <- 2009
df$YEAR[df$SHYEAR =='2009-10'] <- 2010
df$YEAR[df$SHYEAR =='2010-11'] <- 2011
df$YEAR[df$SHYEAR =='2011-12'] <- 2012
df$YEAR[df$SHYEAR =='2012-13'] <- 2013
df$YEAR[df$SHYEAR =='2013-14'] <- 2014

df <- subset(df, df$YEAR>2003)

barplot3 <- ggplot(data=df,aes(x=factor(YEAR))) +
    geom_bar(fill="limegreen") +
    geom_text(stat='count',aes(label=..count..),vjust=-2, size=3.5) +
    theme_minimal()+
    theme(axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    ggtitle("PLOT 3 - New Non-Charter Districts per year 2004-2014")

ggplotly(barplot3)

```
<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# common map properties
g <- list(
  scope = 'usa',
  showland = T,
  landcolor = toRGB("gray90"),
  showcountries = F,
  subunitcolor = toRGB("white")
)

one_map <- function(dat) {
  plot_geo(dat) %>%
    add_markers(x = ~LON, y = ~LAT, color = I("forestgreen"), alpha = 0.5) %>%
    add_text(x = -78, y = 47, text = ~unique(YEAR), color = I("black")) %>%
    layout(geo = g)
}

df %>%
  group_by(YEAR) %>%
  do(map = one_map(.)) %>%
  subplot(nrows = 4) %>%
  layout(
    showlegend = FALSE,
    title = 'New Non-Charter Districts per year 2004-2014<br> Source: <a href="https://nces.ed.gov/ccd/pubagency.asp">NCES</a>',
    width = 1000,
    height = 900,
    hovermode = FALSE
  )

```
